Parallel Information Set Generation for Kriegspiel

### 2012-07-15 sbac/paper.tex
table captions - removed centering
Updates and error fixes to ISG section; a few caption updates

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
that poin. At move 9, Black attempts to block a
potential threat along the diagonal by moving a pawn to g6, then attempts to
escape or capture a threat at f7.  When both of these attempts fail, Black can
infer that his king is being threatened by a protected bishop or queen at f7.
The attempt to move to e7 is Black's way of finding out whether the threatening
piece is a queen or bishop.  When this move fails, black knows it is White's
queen and ultimately retreats to d8.  Note that the failed moves by White on
turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\caption{Abbreviated output from example game.  Three of 25 sequences that are consistent with Black's observations after five moves by each player.  The 25 sequences (or the states that result from them) compose Black's information set at that point in the game. From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}

@@

\label{speedups}
\end{table}

@@

In our next round of experiments, we set out to compare load-balancing
strategies for the problem.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently---if only marginally---better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is probably also related to the
fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}

\caption{Effect of load balancing strategy}
\label{3schemes}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\caption{Grainsize Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\end{figure*}
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
\label{3depths}
\end{figure}

@@

\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Kriegspiel position from 2 games, position 10}
\label{prob4} 
\end{figure}

@@

\label{tab:prob4}
\end{table}

@@

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}
\label{tab:prob5}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\label{bestspeedups}
\end{table}

>>>>

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\vspace{-0.2in}
\end{figure}

@@

Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
that poin. At move 9, (Figure~\ref{filteredlisting}), Black attempts to block a
potential threat along the diagonal by moving a pawn to g6, then attempts to
escape or capture a threat at f7.  When both of these attempts fail, Black can
infer that his king is being threatened by a protected bishop or queen at f7.
The attempt to move to e7 is Black's way of finding out whether the threatening
piece is a queen or bishop.  When this move fails, black knows it is White's
queen and ultimately retreats to d8.  Note that the failed moves by White on
turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\vspace{-0.2in}
\caption{Abbreviated output from example game.  Three of 25 sequences that are consistent with Black's observations after five moves by each player.  The 25 sequences (or the states that result from them) compose Black's information set at that point in the game. From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\vspace{-0.3in}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}

@@

\label{speedups}
\vspace{-0.3in}
\end{table}

@@

In our next round of experiments, we set out to compare load-balancing
strategies for the problem.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently---if only marginally---better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is probably also related to the
fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
\vspace{-0.1in}

\caption{Effect of load balancing strategy}
\label{3schemes}
\vspace{-0.2in}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\vspace{-0.1in}
\caption{Grainsize Control for different problem sizes}
%\vspace{-0.1in}
\label{fig:dep}
\vspace{-0.1in}
\end{figure*}
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time, to help see the isoefficiency.}
\vspace{-0.2in}
\label{3depths}
\end{figure}

@@

\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Kriegspiel position from 2 games, position 10}
\label{prob4} 
\vspace{-0.2in}
\end{figure}

@@

\vspace{-0.2in}
\label{tab:prob4}
\end{table}

@@

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}

\vspace{-0.2in}
\label{tab:prob5}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}
